### 2015-05-04 paper/wish/wsc15paper.tex
conference

%**************************************************************************
%*
%*  Paper: ``INSTRUCTIONS FOR AUTHORS OF LATEX DOCUMENTS''
%*
%*  Publication: 2015 Winter Simulation Conference Author Kit
%*
%*  Filename: wsc15paper.tex
%*
%*  Date: January 31, 2001   Time:  9:45 PM
%*      BASE of current version: Feb 01, 2010 (primary WSC'10 LaTeX file)
%*
%*  Word Processing System: TeXnicCenter and MiKTeX
%*
%*
%*  All files need the following
\input{wsc15style.tex}     % download from author kit.  Style files for wsc formatting. Don't remove this line - required for generating the final paper!

\documentclass{wscpaperproc}
\usepackage{latexsym}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{mathptmx}

%
%****************************************************************************
% AUTHOR: You may want to use some of these packages. (Optional)
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
%****************************************************************************

%
%****************************************************************************
% AUTHOR: If you do not wish to use hyperlinks, then just comment
% out the hyperref usepackage commands below.

%% This version of the command is used if you use pdflatex. In this case you
%% cannot use ps or eps files for graphics, but pdf, jpeg, png etc are fine.

\usepackage[pdftex,colorlinks=true,urlcolor=blue,citecolor=black,anchorcolor=black,linkcolor=black]{hyperref}

%% The next versions of the hyperref command are used if you adopt the
%% outdated latex-dvips-ps2pdf route in generating your pdf file. In
%% this case you can use ps or eps files for graphics, but not pdf, jpeg, png etc.
%% However, the final pdf file should embed all fonts required which means that you have to use file
%% formats which can embed fonts. Please note that the final PDF file will not be generated on your computer!
%% If you are using WinEdt or PCTeX, then use the following. If you are using
%% Y&Y TeX then replace "dvips" with "dvipsone"

%%\usepackage[dvips,colorlinks=true,urlcolor=blue,citecolor=black,%
%% anchorcolor=black,linkcolor=black]{hyperref}
%****************************************************************************

%
%****************************************************************************
%*
%* AUTHOR: YOUR CALL!  Document-specific macros can come here.
%*
%****************************************************************************

% Private macros here (check that there is no clash with the style)
\newcommand{\lambdahat}{\widehat{\lambda}}
\newcommand{\xtilde}{\tilde{x}}
\newcommand{\qhat}{\widehat{q}}
\newcommand{\as}{\ a.s.}
\newcommand{\zap}[1]{}
\newcommand{\Ncal}{\mathcal{N}}
\newcommand{\sigmahat}{\hat{\sigma}}
\newcommand{\Nat}{\mathbb{N}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\Pb}[2]{\mathbb{P}_{#1}\left\{#2\right\}}
\newcommand{\Ybar}{\overline{Y}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
% \newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
% \newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\g}{\,\vert\,}
\newcommand{\ind}[1]{1_{\left\{#1\right\}}}
\newcommand{\Fcal}{\mathcal{F}}
\newcommand{\Gcal}{\mathcal{G}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\xhat}{\hat{x}}
\newcommand{\xwig}{\tilde{x}}
\newcommand{\PZ}{\mbox{PZ}}
\newcommand{\PCS}{\mbox{PCS}}
\newcommand{\PGS}{\mbox{PGS}}
\newcommand{\uLFC}{\underline{u}}
\newcommand{\e}[1]{\left\{ #1 \right\}}
\newcommand{\T}{\mathbb{T}} % Time index set.
\newcommand{\CS}{\mbox{CS}}
\newcommand{\Ft}{\mathcal{F}_t}
\newcommand{\F}[1]{\mathcal{F}_{#1}}
\newcommand{\Ftau}{\mathcal{F}_\tau}
\newcommand{\Rcal}{\mathcal{R}}
\newcommand{\upthresh}{P}
\newcommand{\xstar}{X^*}
\newcommand{\cmax}{1-(P^*)^{\frac1{k-1}}}
\newcommand{\NewN}{M}
\newcommand{\thetavec}{\vec{\theta}}
\newcommand{\uvec}{\vec{u}}
\newcommand{\sigmavec}{\lambda}
\newcommand{\sigmacom}{\sigma}
\newcommand{\sigmascal}{\lambda}
\newcommand{\PstarB}{\mathcal{P}^*_B}
\newcommand{\Ymod}{Y'}
\newcommand{\ceil}{\mathrm{ceil}}
% How to state the common variance assumption at the beginning of lemmas, propositions and theorems
\newcommand{\homog}{Suppose $\lambda^2_x = \sigma^2>0\ \forall x$.  }
\newcommand{\algref}[1]{Alg.~\ref{#1}}

% If you use theoremes
\newtheoremstyle{wsc}% hnamei
{3pt}% hSpace abovei
{3pt}% hSpace belowi
{}% hBody fonti
{}% hIndent amounti1
{\bf}% hTheorem head fontbf
{}% hPunctuation after theorem headi
{.5em}% hSpace after theorem headi2
{}% hTheorem head spec (can be left empty, meaning `normal')i

\theoremstyle{wsc}
\newtheorem{theorem}{Theorem}
\renewcommand{\thetheorem}{ \arabic{theorem}}
\newtheorem{corollary}{Corollary}
\renewcommand{\thecorollary}{\arabic{corollary}}
\newtheorem{definition}{Definition}
\renewcommand{\thedefinition}{\arabic{definition}}
\newtheorem{lemma}{Lemma}
\renewcommand{\thelemma}{\arabic{lemma}}

%#########################################################
%*
%*  The Document.
%*
\begin{document}

%***************************************************************************
% AUTHOR: AUTHOR NAMES GO HERE
% FORMAT AUTHORS NAMES Like: Author1, Author2 and Author3 (last names)
%
%		You need to change the author listing below!
%               Please list ALL authors using last name only, separate by a comma except
%               for the last author, separate with "and"
%
\WSCpagesetup{Toscano-Palmerin, Frazier}

% AUTHOR: Enter the title, all letters in upper case
\title{ASYMPTOTIC VALIDITY OF A FULLY SEQUENTIAL ELIMINATION PROCEDURE FOR INDIFFERENCE-ZONE RANKING AND SELECTION WITH TIGHT BOUNDS ON PROBABILITY OF CORRECT SELECTION}

\maketitle

\begin{figure*}[htb]
{
\centering
Saul Toscano-Palmerin \\
Peter I. Frazier \\
\vspace{12pt}
Cornell University \\
257 Rhodes Hall \\
206 Rhodes Hall \\
Ithaca, NY 14853, USA\\
}
\end{figure*}

\section*{ABSTRACT}
We consider the indifference-zone (IZ) formulation of the ranking
and selection problem in which the goal is to choose with high probability 
an alternative with similar mean than the one with largest mean. 
Conservatism leads classical IZ procedures
to take too many samples in problems with many alternatives. The Bayes-inspired
Indifference Zone (BIZ) procedure, proposed in \citeN{Frazier:BIZ}, is
less conservative than previous procedures, but its proof of validity
requires strong assumptions, specifically it assumes that the variances are known 
and have an integer multiple structure. In this paper, we consider a slightly
modification of the original BIZ procedure in order to simplify the analysis, and we
present a new proof of its asymptotic validity that relaxes these assumptions.
Specifically, we prove the validity of the algorithm when the variances are known
and the difference between the best alternative and the second best goes to zero.

\section{INTRODUCTION}
\label{sec:intro}

There are many applications where we have to choose the best alternative
among a finite number of simulated alternatives. For example, in inventory problems, we may want to choose the best inventory policy $(s,S)$ for a finite number of values of $s$ and $S$. The ranking and selection problem has to decide how many samples are needed to find the alternative with the largest mean. A good solution to the problem is efficient and accurate, i.e. the procedure proposed for the problem must balance between the number of samples required and quality of the selection.

This paper considers the indifference-zone (IZ) formulation of
the ranking and selection problem, in which the distance between the best system and
the other systems is sufficiently large and the best system is chosen with probability larger
than some threshold given by the user. This property is called the IZ guarantee, and the preference zone (PZ)
is defined as the set of system configurations under which the difference between the best system and the others is
at least some given  $\delta>0$. The paper \citeN{Bechhofer:1954} is considered the seminal work, and early work is 
presented in the monograph \citeN{Bechhofer:1968}. Some compilations of the theory developed in the area can be found
in \citeN{Bechhofer:1995}, \citeN{swisher:survey}, \citeN{kim:20062} and \citeN{kim:2007}. Beyond these classical approaches, 
there are the Bayesian approach \cite{Frazier:Tutorial} and the optimal computing budget allocation approach \cite{chen:ocva}.

A good IZ procedure satisfies the IZ guarantee and requires as few samples as possible.
The first IZ procedures presented in \citeN{Bechhofer:1954}, \citeN{Paulson:01}, \citeN{Fabian}, \citeN{Rinott},
\citeN{Harmann:1988}, \citeN{hartmann:sequential}, \citeN{Paulson:1994} 
satisfy the IZ guarantee, but they usually take too many samples when there are many alternatives,
in part because their probability of correct selection (PCS) is much
larger than the probability specified by the user \cite{kim:Conserv}. One reason for this
is that these procedures use Bonferroni's inequality, which leads then to sample more
than necessary. More recently, new algorithms were developed in \citeN{Kim:2001}, \citeN{Goldsman:2002}, 
\citeN{Hong}, and they improve the performance but they still
use the Bonferroni's inequality, and so the methods are inefficient
when there are many alternatives. Procedures in \citeN{kim:2011},
\citeN{dieker:2012} do not use the Bonferroni's inequality 
when there are only three alternatives, but when they compare more than three
alternatives, they do use the Bonferroni's inequality.

Conservative algorithms lead to take more samples than needed, and consequently conservative procedures are unpopular 
when there are more than a few hundred of alternatives. Three common sources of conservativeness are
 the Bonferroni's inequality, the change from discrete process to continuous process, and bound the worst case which is not a common case.
  \citeN{Frazier:BIZ} eliminates one source of conservativeness: the Bonferroni's inequality. He presented  a new sequential elimination
IZ procedure, called BIZ (Bayes-inspired Indifference Zone), whose
lower bound on worst-case probability of correct selection in the
preference zone is tight in continuous time, and almost tight in the
discrete time. In numerical experiments, the number of samples required
by BIZ is significantly smaller than that of procedures like the KN procedure of \citeN{Kim:2001} and
the $\mathcal{P}_{B}^{*}$ procedure of \citeN{Bechhofer:1968}, 
especially on problems with many alternatives. Unfortunately, the
proof that the BIZ procedure satisfies the IZ guarantee for the discrete-time
case assumes that variances are known and have an integer multiple
structure which is not very realistic.

The contribution of this work is that we prove the asymptotic validity of the 
BIZ procedure for the discrete-time case when the variances are known and $\delta$ goes to zero. 
In order to simplify analysis, we consider a slightly modification of this procedure.
Although we only proved it in the known variance setting, it is still a significantly
generalization over the assumption that the variances are known and have an integer multiple
structure. Furthermore, we do not need to assume that the alternatives follow a normal distribution. 
The only assumptios are that the alternatives are independent, identically distributed and have
finite variance. We conjecture that these techniques can be used to show the validity of this procedure when
the variances are unknown, and we present numerical experiments that support this belief. 
\citeN{kim:asympt} also proves the asymptotical validity of a IZ procedure in the same limit.
Our proof shares some similarities since we both use a central limit theorem.

This paper is organized as follows: In $\text{ß}2$, we recall the
indifference-zone ranking and selection problem. In $\text{ß}3$,
we recall the Bayes-inspired IZ (BIZ) procedure from \citeN{Frazier:BIZ}. In $\text{ß}4$,
we present the proof of the validity of the algorithm when the variances
are known. In $\text{ß}5$, we present some numerical experiments. 
In $\text{ß}6$, we conclude.

\section{INDIFFERENCE-ZONE RANKING AND SELECTION}

Ranking and Selection is a problem where we have to select the best system
among a finite set of alternatives, i.e. the system with the largest
mean. The method selects a system as the best based on the samples
that are observed sequentially over time. We suppose that samples
are identically distributed and independent, over time and across alternatives, and
each alernative $x$ has mean $\mu_{x}$. We define $\mu=(\mu_{1},\ldots,\mu_{k})$.

If the best system is selected, we say that the procedure has made the \emph{correct selection}
(CS). We define the \emph{probability of correct selection} as 
\[
\mbox{PCS}\left(\mu\right)=\mathbb{P}_{\mu}\left(\hat{x}\in\mbox{arg max}_{x}\mu_{x}\right)
\]
where $\hat{x}$ is the alternative chosen by the procedure and $\mathbb{P}_{\mu}$
is the probability measure under which samples from system $x$ have
mean $\mu_{x}$ and finite variance $\lambda_{x}^{2}$.

In the Indifference-Zone Ranking and Selection, the procedure is indifferent
in the selection of a system whenever the means of the populations
are nearly the same. Formally, let $\mu=\left[\mu_{k},\ldots,\mu_{1}\right]$
be the vector of the true means, the \emph{indifference zone} is defined
as the set $\left\{ \mu\in\mathbb{R}^{k}:\mu_{\left[k\right]}-\mu_{\left[k-1\right]}<\delta\right\} $.
The complement of the indifference zone is called the \emph{preference
zone} (PZ) and $\delta>0$ is called the indifference zone parameter.
We say that a procedure meets the \emph{indifference-zone (IZ) guarantee
}at $P^{*}\in\left(1/k,1\right)$ and $\delta>0$ if
\[
\mbox{PCS}\left(\mu\right)\geq P^{*}\mbox{ for all }\mu\in\mbox{PZ}\left(\delta\right).
\]
We assume $P^{*}>1/k$ because IZ guarantees can be meet by choosing
$\hat{x}$ uniformly at random from among $\left\{ 1,\ldots,k\right\} $.

\section{THE BAYES-INSPIRED IZ (BIZ) PROCEDURE}

BIZ is an elimination procedure. This procedure maintains a set of
alternatives that are candidates for the best system, and it takes samples from each
alternative in this set at each point in time. At beginning, all alternatives
are possible candidates for the best system, and over the time alternatives are eliminated.
The procedure ends when there is only one alternative in the contention
set and this remain alternative is chosen as the best.

Frazier \cite{Frazier:BIZ} showed that the BIZ procedure with known common variance
satisfies the IZ guarantee when the systems follow the normal distribution
, with tight bounds on worst-case preference-zone in continuous time.
He also proved that this procedure retains the IZ guarantee when the
systems follow the normal distribution, and the variances are known
and are integer multiples of a common value. The continuous time version
of this procedure also satisfies the IZ guarantee, with a tight worst-case
preference-zone PCS bound.

The discrete-time BIZ procedure for unknown and/or heterogeneous sampling
variances is given below.  It takes a variable number of samples
from each alternative, and $n_{tx}$ is this number. This algorithm
depends on a collection of integers $B_{1},\ldots,B_{k}$, $P^{*},c,\delta$
and $n_{0}$. $n_{0}$ is the number of samples to use in the first
stage of samples, and $100$ is the recommended value for $n_{0}$.
$B_{x}$ controls the number of samples taken from system $x$ in
each stage. The procedure presented is a slightly modification of the original BIZ 
procedure where $z \in \mbox{arg max}_{x\in A} \lambdahat^2_{tx}$, instead 
of $z \in \mbox{arg min}_{x\in A} n_{tx} / \lambdahat^2_{tx}$.

For each $t$, $x\in\left\{ 1,\ldots,k\right\} $, and subset $A\subset\left\{ 1,\ldots,k\right\} $,
we define a function
\[
q'_{tx}\left(A\right)=\mbox{exp}\left(\delta\beta_{t}\frac{Z_{tx}}{n_{tx}}\right)\left/\sum_{x'\in A}\mbox{exp}\left(\delta\beta_{t}\frac{Z_{tx'}}{n_{tx'}}\right),\right.\mbox{ }\beta_{t}=\frac{\sum_{x'\in A}n_{tx'}}{\sum_{x'\in A}\hat{\lambda}_{tx'}^{2}}
\]
where $\hat{\lambda}_{tx'}^{2}$ is the sample variance of all samples
from alternative $x$ thus far and $Z_{tx}=Y_{n_{tx},x}$.

\vspace{5mm}

\paragraph{Algorithm: Discrete-time implementation of BIZ, for unknown and/or heterogeneous variances.}    
\begin{algorithmic}[1]   
\label{alg:hetero-BIZ}   
\REQUIRE $c \in [0,\cmax]$, $\delta>0$, $P^*\in(1/k,1)$, $n_0\ge0$ an integer, $B_1,\ldots,B_k$ strictly positive integers.  Recommended choices are $c=\cmax$, $B_1=\cdots=B_k=1$ and $n_0$ between $10$ and $30$.     If the sampling variances $\lambda^2_x$ are known, replace the estimators     
$\lambdahat^2_{tx}$ with the true values $\lambda^2_x$, and set $n_0=0$.

\STATE For each $x$, sample alternative $x$ $n_0$ times and set $n_{0x} \leftarrow n_0$.     
Let $W_{0x}$ and $\lambdahat^2_{0x}$ be the sample mean and sample variance respectively of these samples.     Let $t\leftarrow 0$.Let $z \in \mbox{arg max}_{x\in A} \lambdahat^2_{tx}$.     
\STATE Let $A \leftarrow \{ 1,\ldots, k\}$, $\upthresh \leftarrow P^*$.
\WHILE{$x\in\mbox{max}_{x\in A} q'_{tx}\left(A\right)<P$}
\WHILE{$\mbox{min}_{x\in A} q'_{tx}\left(A\right) \le c$}
 \STATE Let $x\in\mbox{arg min}_{x\in A} q_{tx}\left(A\right)$.
    \STATE Let $\upthresh \leftarrow \upthresh/(1-q_{tx}\left(A\right))$.     
\STATE Remove $x$ from $A$.
\ENDWHILE    
\STATE For each $x\in A$, let      $n_{t+1,x} = \ceil\left( \lambdahat^2_{tx} (n_{tz} + B_z) / \lambdahat^2_{tz} \right)$.     \STATE For each $x\in A$, if $n_{t+1,x}>n_{tx}$, take $n_{t+1,x}-n_{tx}$ additional samples from alternative $x$.  Let $W_{t+1,x}$ and $\lambdahat^2_{t+1,x}$ be the sample mean and sample variance respectively of all samples from alternative $x$ thus far.    
\STATE Increment $t$.
 \ENDWHILE
  \STATE Select $\xhat \in\mbox{arg max}_{x\in A} Z_{tx} / n_{tx}$ as our estimate of the best.

\end{algorithmic}

\vspace{10mm}
This algorithm generalizes the BIZ procedure with known common variance. In that case, we have that$B_1=\cdots=B_k=1$ and $n_{tx}=t$. The algorithm 2 can be generalized to the continuous case (See  \citeN{Frazier:BIZ}).

\section{ASYMPTOTIC VALIDITY WHEN THE VARIANCES ARE KNOWN}

In this section we prove that the BIZ procedure satisfies asymptotically the IZ guarantee
when the variances are known. This means that we consider a collection of
ranking and selection problems parametrized by $\delta>0$.  
For the problem given $\delta$, we suppose that the vector of the true means
$\mu=\left[\mu_{k},\ldots,\mu_{1}\right]$ is equal to $\delta a$ for some fixed $a\in\mathbb{R}^{k}$ that
does not depend on $\delta$ and $a_{k}>a_{k-1}\geq \cdots \geq a_{1}$, $a_{k}-a_{k-1}>1$.
Moreover, the variances of the alternatives are finite, strictly greater than zero and do not depend on $\delta$.
We also suppose that samples from system $x\in\left\{ 1\ldots,k\right\} $ are identically distributed
and independent, over time and across alternatives. We also define
$\lambda_{z}^{2}:=\max_{i\in\left\{ 1\ldots,k\right\} }\lambda_{i}^{2}$.

Any ranking and selection algorithm can be viewed as mapping from
paths of the $k$-dimensional discrete-time random walk $\left(Y_{tx}:t\in\mathbb{N},x\in\left\{ 1,\ldots,k\right\} \right)$
onto selection decisions. The BIZ procedure's mapping from paths onto
selections decisions can be understood as the composition of three
simpler maps.

The first is the mapping from the raw discrete-time random walk $\left(Y_{tx}:t\in\mathbb{N},x\in\left\{ 1,\ldots,k\right\} \right)$
onto a time changed version of this random walk, written as $\left(Z_{tx}:t\in\mathbb{N},x\in\left\{ 1,\ldots,k\right\} \right)$,
where we recall $Z_{tx}=Y_{n_{x}\left(t\right),t}$ is the sum of
the samples from alternative $x$ observed by stage $t$.

The second maps this time-changed random walk through a non-linear
mapping for each $t,x$ and subset $A\subset\left\{ 1,\ldots,k\right\} $,
to obtain $\left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right)$,
where

\[
q'_{tx}\left(A\right)=\mbox{exp}\left(\delta\beta_{t}\frac{Z_{tx}}{n_{tx}}\right)\left/\sum_{x'\in A}\mbox{exp}\left(\delta\beta_{t}\frac{Z_{tx}}{n_{tx}}\right)\right.:=q'\left(\left(Z_{tx},t:x\in A\right),\delta,t\right)
\]
where we note that $n_{x}\left(t\right)$ and $\beta_{t}$ are deterministic
in the version of the known-variance BIZ procedure that we consider
here.

The third maps the paths of $\left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right)$
onto selection decisions. Specifically, this mapping finds the first
time that $\mbox{max}_{x\in A}q_{tx}\left(A\right)$ is larger than
some threshold $P$, and then it chooses the system with the largest
empirical mean $Z_{tx}/n_{tx}$.

Later, we can approximate the previous process by the continuous time
process
\begin{eqnarray}
q_{tx}\left(A\right) & = & q\left(\left(Z_{\frac{t}{\delta^{2}}x},\frac{t}{\delta^{2}}:x\in A\right),\delta,A\subset\left\{ 1,\ldots,k\right\} \right)\label{eq:0.001-1}
\end{eqnarray}
where $t\geq0$ . Then we will show the difference in selection decisions
vanishes as $\delta\rightarrow0$.

After that, we would like to use a kind of central limit theorem for
$Z_{\frac{t}{\delta^{2}}x}$ when $\delta$ goes to zero to work with
a Brownian motion instead. In order to do this, we are going to rewrite
equation (1) in terms of a centralized version of $Z$:

\[
\mathcal{C}_{x}\left(\delta,t\right)=\frac{Y_{n_{x}\left(t\right),x}-t\lambda_{x}^{2}\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}\delta}}
\]
and the resulting expression is

\[
q_{tx}\left(A\right)=q\left(\left(C_{x}\left(\delta,t\right)\frac{\lambda_{x}^{2}}{\delta\lambda_{z}^{2}}+\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(n_{0}+\frac{t}{\delta^{2}}\right)\delta a_{x},\frac{t}{\delta^{2}}:x\in A\right),\delta,A\subset\left\{ 1,\ldots,k\right\} \right).
\]

Thus, we can see the selection decision resulting from the BIZ algorithm
as the result of applying some mapping $f\left(\cdot,\delta\right)$
to the paths $t\mapsto\mathcal{C}_{x}\left(\delta,t\right)$ .

The last part of our proof is to observe that (1) $\mathcal{C}\left(\delta,\cdot\right)$
converges to a multivariate Brownian motion $W$ as $\delta$ goes
to 0; (2) the function $f$ has a continuity property that causes
\[
f\left(\mathcal{C}\left(\delta,\cdot\right),\delta\right)\Rightarrow g\left(W\right)
\]
where g is the selection decision from applying the BIZ procedure
in continuous time; and (3) the BIZ procedure satisfies the IZ guarantee
when applied in continuous time.

First, we are going to see that the the standardized sum of the output
data converges to a Brownian motion in the sense of $D_{\infty}:=D[0,\infty)$,
which is the set of functions from $\left[0,\infty\right)$ to $\mathbb{R}$
that are right-continuous and have left-hand limits, with the Skorohod
topology. The definition and the properties of this topology may be
found in Chapter 3 of \citeN{billingsley:convergnce2}.

We briefly recall the definition of convergence of random paths in
the sense of $D_{\infty}$. Suppose that we have a sequence of random
paths $\left(\mathcal{X}_{n}\right)_{n\geq0}^{\infty}$ such that
$\mathcal{X}_{n}:\varOmega\rightarrow D_{\infty}$ where $\left(\Omega,\mathcal{F},\mathbb{P}\right)$
is our probability space. We say that $\mathcal{X}_{n}\Rightarrow\mathcal{X}_{0}$
in the sense of $D_{\infty}$ if $P_{n}\Rightarrow P_{0}$ where $P_{n}:\mathcal{D}_{\infty}\rightarrow\left[0,1\right]$
are defined as $P_{n}\left[A\right]=\mathbb{P}\left[\mathcal{X}_{n}^{-1}\left(A\right)\right]$
for all $n\geq0$ and $\mathcal{D}_{\infty}$ are the Borel subsets
for the Skorohod topology.

The following lemma shows that the standardized sum of the output
data with $t$ changed by $t/\delta^{2}$ converges to a Brownian
motion in the sense of $D_{\infty}$. 
\\
\begin{lemma}
For each $x\in\left\{ 1\ldots,k\right\} $, we define
\[
\mathcal{C}_{x}\left(\delta,t\right)=\frac{Y_{n_{x}\left(t\right),x}-t\lambda_{x}^{2}\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}\delta}}
\]
where $Y_{n,x}$ is the sum of the first $n$ samples from alternative
$x$. Let $x\in\left\{ 1\ldots,k\right\} $, then
\[
\mathcal{C}_{x}\left(\delta,\cdot\right)\Rightarrow W_{x}\left(\cdot\right)
\]
as $\delta\rightarrow0$ in the sense of $D[0,\infty)$, where $W_{x}$
is a standard Brownian motion.

\end{lemma}

\begin{proof}
By Theorem 19.1 of  \citeN{billingsley:convergnce2},

\[
\frac{Y_{n_{x}\left(t\right),x}-\mbox{floor}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right)\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}\Rightarrow W_{x}\left(\cdot\right)
\]
in the sense of $D[0,\infty)$.

Fix $w\in\Omega$. Observe that
\[
\frac{Y_{\mbox{floor}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right),x}-\mbox{floor}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right)\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}-\frac{Y_{\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right),x}-\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right)\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}\rightarrow0
\]
uniformly in $\left[0,s\right]$ for all $s\geq0$ and then by Theorem
A.2 
\[
\frac{Y_{\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right),x}-\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right)\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}\Rightarrow W_{x}\left(\cdot\right)
\]
in the sense of $D[0,\infty)$.

Since $\frac{\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}t\frac{1}{\delta^{2}}-ceil\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}t\frac{1}{\delta^{2}}\right)}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}\rightarrow0$
uniformly on $[0,s]$ for every $s\geq0$, then by Theorem A.2

\[
\frac{Y_{\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right),x}-\mbox{}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right)\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}\Rightarrow W_{x}\left(\cdot\right).
\]

Finally, observe that for fixed $\omega\in\Omega$,
\begin{eqnarray*}
 &  & \frac{Y_{\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right),x}-\mbox{}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right)\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}-\frac{Y_{\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)+n_{0}\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\right),x}-\mbox{}\left(n_{0}\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}+\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right)\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}\\
 & = & \frac{Y_{\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)\right),x}-Y_{\mbox{ceil}\left(\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(\cdot\frac{1}{\delta^{2}}\right)+n_{0}\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\right),x}+\left(n_{0}\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\right)\mu_{x}}{\frac{\lambda_{x}^{2}}{\lambda_{z}}\sqrt{\frac{1}{\delta^{2}}}}\\
 & \rightarrow & 0
\end{eqnarray*}
uniformly in $\left[0,t\right]$ for all $t\geq0$, and so by Theorem
A.2 the result follows.

\end{proof}

Now, we use the product topology in $D^{k}\left[0,\infty\right)$
for $k\in\mathbb{N}$. This topology may be described as the one under
which $\left(Z_{n}^{1},\ldots,Z_{n}^{k}\right)\rightarrow\left(Z_{0}^{1},\ldots,Z_{0}^{k}\right)$
if and only if $Z_{n}^{i}\rightarrow Z_{0}^{i}$ for all $i\in\left\{ 1,\ldots,k\right\} $.
See the Miscellany of  \citeN{billingsley:convergence}. The following corollary follows
from the previous result and independence.

\vspace{5mm}

\begin{corollary}
We have that
\[
\mathcal{C}\left(\delta,\cdot\right):=\left(\mathcal{C}_{x}\left(\delta,\cdot\right)\right)_{x\in A}\Rightarrow W\left(\cdot\right):=\left(W_{x}\left(\cdot\right)\right)_{x\in A}
\]
as $\delta\rightarrow0$ in the sense of $D_{\infty}^{k}$.
\end{corollary}

\vspace{5mm}

Observe that $\mathcal{C}\left(\delta,\cdot\right)\in D\left[0,\infty\right)^{k}$
and the convergence of Lemma 1 is in $D\left[0,\infty\right)^{k}$,
consequently the following natural step is to generalize the continuous-time
procedure for the heterogeneous variance setting, see Appendix B,
to the space $D\left[0,\infty\right)^{k}$. Specifically, for each
$F\in D\left[0,\infty\right)^{k}$, we define $q_{tx}^{F,\delta}\left(A\right)$
as
\[
q_{tx}^{F,\delta}\left(A\right)=q'\left(\left(F_{x}\left(t\right)\frac{\lambda_{x}^{2}}{\delta\lambda_{z}^{2}}+\frac{\lambda_{x}^{2}}{\lambda_{z}^{2}}\left(n_{0}+\frac{t}{\delta^{2}}\right)\delta a_{x},\frac{t}{\delta^{2}}:x\in A\right),\delta,A\subset\left\{ 1,\ldots,k\right\} \right).
\]

Note that if we replace $F$ by $\mathcal{C}\left(\delta,t\right)$,
we get $q{}_{tx}\left(A\right)$ in the previous equation. Using the
continuous-time procedure for the heterogeneous variance setting with
the paths $\left(q_{tx}^{F,\delta}\left(A\right):t\geq0\right)$,
we see that there exists a selection decision for each $F\in D\left[0,\infty\right)^{k}$.
Since we are interested in the right selection, ( i.e. in choosing
$k$) for each $F\in D\left[0,\infty\right)^{k}$, we define the following
function 
\begin{eqnarray*}
f\left(F,\delta\right) & = & \begin{cases}
1 & \mbox{if }k\mbox{ is chosen}\mbox{ }\\
0 & \mbox{otherwise}
\end{cases}
\end{eqnarray*}
 Observe that $f$ is a function that depends in the path $\left\{ q_{tx}^{F,\delta}\left(A\right):t\geq0\right\} $.
We are doing all this because we want to see that $f\left(\mathcal{C}\left(\delta,\cdot\right),\delta\right)$
looks like the right or wrong selection decision of an algorithm that
depends on a standard Brownian motion instead of $\mathcal{C}\left(\delta,\cdot\right)$,
and this will give us the proof because we will be able to use the
theory developed by Frazier (2014) to prove the theorem. Thus, the
following step is to find the limit of $q_{tx}^{F,\delta}\left(A\right)$
when $\delta$ goes to zero, which is called $q_{tx}^{F}\left(A\right)$
and equal to

\[
q_{tx}^{F}\left(A\right):=\mbox{exp}\left(\frac{F_{x}\left(t\right)}{\lambda_{z}}+\frac{1}{\lambda_{z}^{2}}ta_{x}\right)/\sum_{x^{'}\in A}\mbox{exp}\left(\frac{F_{x'}\left(t\right)}{\lambda_{z}}+\frac{1}{\lambda_{z}^{2}}ta_{x^{'}}\right).
\]
 Using the same previous idea, for each $F\in D\left[0,\infty\right)^{k}$,
we define the new functions
\[
\]
\begin{eqnarray*}
g\left(F\right) & = & \begin{cases}
1 & \mbox{if }k\mbox{ is chosen}\mbox{ }\\
0 & \mbox{otherwise}
\end{cases}.
\end{eqnarray*}

So, we want to prove that 
\[
f\left(\mathcal{C}\left(\delta,\cdot\right),\delta\right)\Rightarrow g\left(W\right)
\]
as $\delta\rightarrow0$ in distribution.

In order to prove this, we first proved a continuity property in Lemma
2, which allows us to use Theorem 5.5 of \citeN{billingsley:convergence}, which
implies the desired result. A proof of Lemma 2 may be found in a full version which will be
submitted soon to arXiv. 
\\
\begin{lemma}

Let $\left\{ \delta_{n}\right\} \subset\left(0,\infty\right)$ such
that $\delta_{n}\rightarrow0$. If $D_{s}\equiv\{Z\in D\left[0,\infty\right)^{k}:\mbox{ if }\left\{ Z_{n}\right\} \subset D\left[0,\infty\right)^{k}\mbox{ and }$
$\mbox{lim}{}_{n}d_{\infty}\left(Z_{n},Z\right)=0$ , then the sequence
$\left\{ f\left(Z_{n},\delta_{n}\right)\right\} $ converges to $\left\{ g\left(Z\right)\right\} $$\left.\right\} $$ $,
then $ $$\mathbb{P}\left(W\text{\ensuremath{\in}}D_{s}\right)=1$.

\end{lemma}
\vspace{5mm}
By the extension of the CMT \cite{billingsley:convergence}, we
have the following corollary.
\vspace{5mm}

\begin{corollary}

We have that

\[
f\left(C\left(\delta,t\right),\delta\right)\Rightarrow g\left(W\left(t\right)\right)
\]
in distribution as $\delta\rightarrow0$.

\[
\]
\end{corollary}
\vspace{1mm}

\begin{theorem}
If samples from system $x\in\left\{ 1\ldots,k\right\} $ are identically
distributed and independent, over time and across alternatives, then
$\mbox{lim}_{\delta\rightarrow0}Pr\left\{ \mbox{BIZ selects }k\right\} \geq P*$
provided $\mu_{k}=a_{k}\delta,\mu_{k-1}=a_{k-1}\delta,\ldots,\mu_{1}=a_{1}\delta$,
$a_{k}>a_{k-1}\geq \cdots \geq a_{1}$, $a_{k}-a_{k-1}>1$, and the variances are finite and
do not depend on $\delta$.

Furthermore,
\[
\mbox{inf}_{a\in PZ\left(1\right)}\mbox{lim}_{\delta\rightarrow0}\mathbb{P}\left(CS_{\delta}\right)=P^{*}
\]
where $PZ\left(1\right)=\left\{ a\in\mathbb{R}^{k}:a_{k}-a_{k-1}>1, a_{k}>a_{k-1}\geq \cdots \geq a_{1} \right\} $.

\end{theorem}

\begin{proof}

By the comments given at the beginning of this section, we know that
we can work with the algorithm defined by

\[
q'_{tx}\left(A\right)=q'\left(\left(Z_{\frac{t}{\delta^{2}}x},\frac{t}{\delta^{2}}:x\in A\right),\delta,A\subset\left\{ 1,\ldots,k\right\} \right)
\]
where $t\in\delta^{2}\mathbb{N}$, instead of the algorithm defined
by
\[
q'_{tx}\left(A\right)=q'\left(\left(Z_{tx},t:x\in A\right),\delta,A\subset\left\{ 1,\ldots,k\right\} \right)
\]
where $t\in\mathbb{N}$.

Now, we define
\begin{eqnarray*}
\hat{\tau}_{n+1}\left(\delta\right) & = & \mbox{inf}\left\{ t\in\left\{ \tau_{n}\delta^{2},\left(\tau_{n}+1\right)\delta^{2},\ldots\right\} :\mbox{ min}_{x\in A_{n}}q_{t/\delta^{2},x}^{'}\left(A_{n}\right)\leq c\mbox{ or }\mbox{max}_{x\in A_{n}}q_{t/\delta^{2},x}^{'}\left(A_{n}\right)\geq P_{n}\right\} 
\end{eqnarray*}
We denote the corresponding continuous hitting times by $\left(\tau_{n}\left(\delta\right)\right)_{n}$,
which are defined as

\[
\tau_{n+1}=\mbox{inf}\left\{ t\geq\tau_{n}:\mbox{ min}_{x\in A_{n}}q_{tx}\left(A_{n}\right)\leq c\mbox{ or }\mbox{max}_{x\in A_{n}}q_{tx}\left(A_{n}\right)\geq P_{n}\right\} .
\]

Using that $\mathcal{C}\left(\delta,\cdot\right)$ is right-continuous,
we can prove that $\hat{\tau}_{n}\left(\delta\right)-\tau_{n}\left(\delta\right)\rightarrow0$
with probability $1$ as $\delta\rightarrow0$, and so we can use
$\mathcal{C}\left(\delta,\tau_{n}\left(\delta\right)\right)$ instead
of $\mathcal{C}\left(\delta,\hat{\tau}_{n}\left(\delta\right)\right)$
in the limit.

Let $CS_{\delta}$ be the event of doing a correct selection given
the configuration $\mu_{k}=a_{k}\delta,\mu_{k-1}=a_{k-1}\delta,\ldots,\mu_{1}=a_{1}\delta$
$ $. Then by the previous argument and the Corollary 2, 
\begin{eqnarray*}
\mbox{lim}_{\delta\rightarrow0}\mathbb{P}\left(CS_{\delta}\right) & = & \mbox{lim}_{\delta\rightarrow0}\mathbb{P}\left(f\left(\mathcal{C}\left(\delta,t\right),\delta\right)=1\right)\\
 & = & \mathbb{P}\left(g\left(W\right)=1\right)\\
 & \geq & P^{*}
\end{eqnarray*}
where the last inequality follows from the Theorem 2 of \citeN{Frazier:BIZ}.

Furthermore, by the same theorem 2,
\[
\mbox{inf}_{a\in PZ\left(1\right)}\mathbb{P}\left(g\left(W\right)=1\right)=P^{*}
\]

where $PZ\left(1\right)=\left\{ a\in\mathbb{R}^{k}:a_{k}-a_{k-1}\geq1\right\} $.

\end{proof}

\section{NUMERICAL EXPERIMENTS}

We show the performance of the BIZ procedure in discrete time with maximum elimination
($c=1-(P^*)^\frac{1}{k-1}$)when $\delta$ goes to zero. We plot both the PCS and the expected total number 
of samples simulated divided by the number of systems, denoted by $E[N]/k$. We consider this quotient 
to normalize the number of samples used by the algorithm.

First, we consider the known variance case. Specifically, we consider a slippage configuration 
with $100$ systems normally distributed, where $\mu_{k}=\delta,\mu_{k-1}=0,\dots,\mu_{1}=0$, $\delta$ is 
within the interval $[0.1,10]$, and $\lambda_{100}=1,\lambda_{99}=1+\frac{(0.5)(98)}{99},\cdots,\lambda_{1}=0.5$. Here, $P^*=0.9$ and $n_{0}=0$.
Figure ~\ref{fig: tahi3} shows that in this case the IZ guarantee is always satisfied, and the bound on probability of correct selection is tight when $\delta$ goes to zero.
When $\delta$ is big enough, the PCS is almost one because the difference between the best system and the others is large enough to be easily identifiable by the
BIZ procedure.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plot5}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plot6}
    \end{subfigure}
    \caption{Known variances and $P^*=0.9$. $\lambda_{100}=1,\lambda_{99}=1+\frac{(0.5)(98)}{99},\cdots,\lambda_{1}=0.5,n_{0}=0$.
    In this example, our theorem is true: the IZ guarantee is always satisfied and the inequality is tight as $\delta$ goes to zero.
    \label{fig: tahi3}}
\end{figure}

We now consider the unknown variance case. Specifically, we consider a slippage configuration 
with $100$ systems normally distributed, where $\mu_{100}=\delta,\mu_{99}=0,\dots,\mu_{1}=0$, $\delta$ is 
within the interval $[0.1,10]$, and $\lambda_{100}=10,\lambda_{99}=\cdots=\lambda_{1}=1$. Here, $P^*=0.9$ and $n_{0}=15$.
Figure ~\ref{fig: tahi2} gives us evidence that our theorem should also be true when the variances are unknown. We should note that
this is a difficult example, and still we have a good performance in most of the cases. However, if we want to always satisfy the IZ guarantee, 
we can just increment the parameter $n_{0}$.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plot3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plot4}
    \end{subfigure}
    \caption{Unknown variances and $P^*=0.9$. $\lambda_{100}=10\sqrt(2),\lambda_{99}=10+\frac{(10)(98)}{99}(\sqrt(2)-1),\cdots,\lambda_{1}=10,n_{0}=15$.
    In this example, our theorem is true: the IZ guarantee is satisfied and the inequality is tight as delta goes to zero. If we want that the IZ guarantee be 
    satisfied in all the cases, we can just increment $n_{0}$.
    \label{fig: tahi2}}
\end{figure}

 Now, we consider a pathological case when the variances are known. Specifically, we consider a slippage configuration 
with $100$ systems normally distributed, where $\mu_{100}=\delta,\mu_{99}=0,\dots,\mu_{1}=0$, $\delta$ is 
within the interval $[0.1,10]$, and $\lambda_{100}=10,\lambda_{99}=\cdots=\lambda_{1}=1$. Here, $P^*=0.9$ and $n_{0}=0$.
This configurations was specially chosen to ilustrate our theorem, and so it is harder than typical configurations.
In fact, in most of the configurations the algorithm always satisfies the IZ guarantee. Furthermore, in practice, we would
have run the algorithm with $n_{0}>0$, and the BIZ algorithm would have worked well even in this pathological example. 
Figure ~\ref{fig: tahi} shows that PCS converges to $0.9$ as $\delta$ goes to zero. Specifically, PCS is equal to $0.904$ when $\delta=0.1$.
This figure also shows that the bound is tight  on probability of correct selection. We should also note that the number of samples required increases very fast. 
Here, the points plotted have central confidence intervals of length at most $0.014$.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plot1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plot2}
    \end{subfigure}
    \caption{Known variances, $P^*=0.9$, $\lambda_{100}=10,\lambda_{99}=\cdots=\lambda_{1}=1,n_{0}=0$. This was a hard example to find
    and it was specially chosen to ilustrate our theorem. In practice, we would choose $n_{0}>0$ and the IZ guarantee will be satisfied. \label{fig: tahi}}
\end{figure}

\section{CONCLUSION}

We have proved the asymptotic validity of the Bayes-inspired Zone
procedure \cite{Frazier:BIZ} when the variances are known, which is a
new sequential elimination procedure. This algorithm is relevant because
it takes fewer samples than other IZ procedures, especially for problems
with large numbers of alternatives. Even though this proof does not
guarantee that the algorithm will work for any sample, we know that
it will work if the alternatives are not very different, which are
the most difficult cases.

\section*{ACKNOWLEDGMENTS}
Peter Frazier and Saul Toscano-Palmerin were partially supported by NSF CAREER CMMI-1254298, AFOSR FA9550-12-1-0200, and AFOSR FA9550-15-1-0038.
Peter Frazier was also partially supported by NSF IIS-1247696 and the Atkinson Center for a Sustainable Future Academic Venture Fund.
Saul Toscano-Palmerin was also partially supported by the Mexican Secretariat of Public Education (SEP).

% Please don't exchange the bibliographystyle style
\bibliographystyle{wsc}
% AUTHOR: Include your bib file here
\bibliography{wsc15paper}

\section*{AUTHOR BIOGRAPHIES}

\noindent {\bf SAUL TOSCANO-PALMERIN} is a Ph.D. student of the School of Operations Research and Information Engineering at Cornell University, Ithaca, NY. His research interest is in simulation optimization, machine learning and sequential decision-making under uncertainty. His email address is \email{st684@cornell.edu}. \\

\noindent {\bf PETER I. FRAZIER} is an Assistant Professor of the School of Operations Research and Information Engineering at Cornell University, Ithaca, NY.  He holds a Ph.D. in operations research and financial engineering industrial engineering from Princeton University. His research interests include optimal learning, sequential decision-making under uncertainty, and machine learning, focusing on applications in simulation optimization, design of experiments, materials science, e-commerce and medicine. He is a member of INFORMS, and SIAM. His e-mail address is \email{pf98@cornell.edu}.\\

\end{document}