### 2015-07-02 paper/wish/correctionsWSC.tex

%% LyX 2.1.2.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=1.5cm,bmargin=1.5cm,lmargin=1.5cm,rmargin=1.5cm}
\usepackage{mathrsfs}
\usepackage{babel}
\begin{document}

\title{Changes to the WSC paper}
\maketitle
\begin{enumerate}
\item Frazier 2014 is now correctly cited in the paper.
\item Page, 3rd paragraph: line 3, lower bound on {*}the{*} worst-case probability;
line 4, in the discrete time -> in discrete time; Line 5, order for
references to KN and $\mathscr{P}_{B}^{*}$ has been switched.
\item Page 2, Last paragraph, - it's now Section 2, 3, 4 instead of 2,3,4. 
\item Page 3, 3rd paragraph, and page 4 - vector mu is now indexed from
1 to k. 
\item Page 3, 4th paragraph, - the word among before \{1, ..., k\} was removed.
\item Page 3, 1st paragraph in Section 3, remain -> remaining 
\item 3rd paragraph in Section 3 -Before: It takes a variable number of
samples from each alternative, and ntx is this number. Now: This algorithm
takes a variable number of samples from alternative $x$ at time $t$,
and $n_{tx}$ is this number (its definition may found in the algorithm
given below). 
\item End of page 3, Before: $Z_{tx}=Y_{n_{tx},x}$ . Now: and $Z_{tx}=Y_{n_{tx},x}$
is the sum of the samples from alternative $x$ observed by stage
$t$.
\item Section 4, 3rd paragraph, Before: The first is the mapping; The second
maps; The third maps. Now: The first mapping is ; The second one maps;
The third one maps
\item Section 6, Zone -> Indifference Zone 
\item References: before: \textquotedblleft automatic control\textquotedblright ;
S.-H. Kim. Now: \textquotedblleft Automatic Control\textquotedblright ;
S. Kim\end{enumerate}

\end{document}